---
title: "Initial Korean covid 19"
author: "Aubrey Garcia"
date: "05/31/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

###introduction 


Data was taken of several individuals regarding age, state, and sex and  who is more susceptible to dying as a result of sex. This study data was taken from  South Korea 1 Jan to 1 May 2020. Furthermore, this leads to the question. Does the sex interfere with the chance of dying?


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

$H_0$: Sex of individual is not an independent  of individuals death.

$H_A$: Sex is an independent to the individuals death.


#methods

Result is a categorical variable, sex is categorical. 
Result depends on sex, we use helper file
CAT~CAT


#Results


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
The barchart shows that males have a high chance of dying than females.




  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that among the people who died, 64% of those who died
were males, while only 36% of the females died. Concluding that a higher percent of the individuals who died were men.

The results of the numerical summary also show that among the 98.7% of females that were alive, only 1.4% died. The graph also shows that among the 97% of males, 3% of them died. Showing that the higher percent of deaths were males. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results, using a chi-squired test, showed that the p-vaule is 0.001, therefore we reject the null hypothesis.
In the second part of the chi-squared test, the p-value is  0.0012, which is also below .5 as so we reject the null hypothesis

## Conclusions

