---
title: "Initial Korean covid 19"
author: "Aubrey Garcia"
date: "05/31/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

###introduction 


Data was taken of several individuals regarding age, state, and sex and  who is more susceptible to dying as a result of sex. This study data was taken from  South Korea 1 Jan to 1 May 2020. Furthermore, this leads to the question. Does the sex interfere with the chance of dying under the conditions of having covid in korea.?


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

$H_0$: Sex of individual is not an independent  of individuals death under the conditions of having covid in korea..

$H_A$: Sex is an independent to the individuals death under the conditions of having covid in korea..


#methods

Result is a categorical variable, sex is categorical. 
Result depends on sex, we use helper file
CAT~CAT


#Results


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
The barchart shows that males have a high chance of dying than females under the conditions of having covid in korea..




  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that among the people who died, 64% of those who died
were males, while only 36% of the females died. Concluding that a higher percent of the individuals who died were men.

The results of the numerical summary also show that among the 98.7% of females that were alive, only 1.4% died. The graph also shows that among the 97% of males, 3% of them died. Both geneder under the conditions of having covid in korea. Showing that the higher percent of deaths were males . 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results, using a chi-squired test, showed that the p-vaule is 0.001, therefore we reject the null hypothesis.
In the second part of the chi-squared test, the p-value is  0.0012, which is also below .5 as so we reject the null hypothesis

The Fisher's Exact Test for Count Data shows that the p-vaule is 0.009, which reject the null hypothesis too. The data also shows that males have a odd ratio of 2.3 likability to die than females under the conditions of having covid in korea.


## Conclusions

To conclude, after being asked if sex in an independent to individuals deaths under the conditions of having covid in korea, results showed that it does. Results from the bar chart also showed the amount of males who died compared to women. Results as taken from the fishers test, showed that males have a odd ratio of 2.3,  more times likely to die than women. This was concluding after the p-vaule taken form the chi-squre test was 0.001 which less than .5, rejecting the null hypothesis. Concluding that sex and death do relate to each other.

